---
title: "Data Analyses"
output: html_document
date: "2023-02-12"
---

# Import
```{r, echo=F, message=F}
# libraries:
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape)
library(glue)
library(mgcv)
library(nlme)
library(Matrix)
library(lme4)
library(wesanderson)
library(viridis)
library(tidyverse)

options(scipen=5, digits = 5)
```

# Prepare the data to analysis 
```{r, message=F, warning=F, error=F}
# Reading the file:
# dir <- "C:/Users/Merav/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_22.2.23.xlsx"
# dir <- "C:/Users/edena/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_22.2.23.xlsx"
# dir<-"C:/Users/Shahar/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_22.2.23.xlsx"
dir <- "C:/Users/shahars/OneDrive - Playtika Ltd/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_22.2.23.xlsx"
df <- read_excel(dir, sheet = 1)
df <- df %>% mutate(DOB = as.Date(DOB)) 

df <- df %>% filter(is.na(lost_of_followup))# Filtering only participants that completed the survey
df <- df %>% filter(month_survey_complete == 2) # Filtering only participants that completed the second year
df <- df[,-c(51:88)]# Removing non relevant columns: from type_of_allergy to 12_month_FA_type_allergy

not_relevant <- c('allergist_followup', 'last_spt', 'epipen', 'mistake_exposure', 'reaction_type24', 'diagnosis_by', 'lost_of_followup',"mother_no_comorbidities","mother_other","father_no_comorbidities","father_other","sibling1_other","sibling1_no_comorbidities","sibling2_other","sibling2_no_comorbidities","sibling3_other","sibling3_no_comorbidities","sibling4_other","sibling4_no_comorbidities")
df <- df[-which(colnames(df) %in% not_relevant)] # Removing columns with alot on NA:

# Adding binary columns 
df$family_Atopic_diseases_sum <- rowSums(df[,14:37]) # Did any one from family (mother, father, sib) have *any* of the disease?
df$mother_Atopic_diseases_sum <- rowSums(df[,14:17]) # Did a disease appear in a specific family member?
df$father_Atopic_diseases_sum <- rowSums(df[,18:21])
df$sibling_Atopic_diseases_sum <- rowSums(df[,22:37])
# df = df %>% mutate(sibling_Atopic_diseases_sum = ifelse(number_of_children ==0, NA, sibling_Atopic_diseases_sum)) # inseret NA to sibiling atopic disease when there is no Sibilings 

# Sum each disease 
df <- df %>% mutate(sibling_asthma_sum = sibling1_asthma + sibling2_asthma + sibling3_asthma + sibling4_asthma,
                    sibling_AD_sum = sibling1_atopic_dermatitis + sibling2_atopic_dermatitis + sibling3_atopic_dermatitis + sibling4_atopic_dermatitis,
           sibling_Allergic_rhinitis_sum = sibling1_Allergic_rhinitis + sibling2_Allergic_rhinitis + sibling3_Allergic_rhinitis + sibling4_Allergic_rhinitis,
                    sibling_FA_sum = sibling1_food_allergy + sibling2_food_allergy + sibling3_food_allergy + sibling4_food_allergy)
              
# Binary variable for each family member 
df <- df %>% mutate(family_Atopic_diseases = ifelse(family_Atopic_diseases_sum >0,1,0),
                    mother_Atopic_diseases = ifelse(mother_Atopic_diseases_sum>0,1,0),
                    father_Atopic_diseases = ifelse(father_Atopic_diseases_sum > 0, 1,0),
                    sibling_Atopic_diseases = ifelse(sibling_Atopic_diseases_sum > 0 ,1, 0))


# Changing to binary variables:
df$gender[df$gender == 2] <- 0
df$`Jew/Arab`[df$`Jew/Arab` == 2] <- 0
df$food_allergy_24_month[df$food_allergy_24_month == 2] <- 0
df$inhalation_24_month[df$inhalation_24_month == 2] <- 0
df$prevention_therapy[df$prevention_therapy == 2] <- 0
df$AD_24_month[df$AD_24_month == 2] <- 0
df$steroids_treatment24[df$steroids_treatment24 == 2] <- 0
df$ABX_treatment_24_months[df$ABX_treatment_24_months == 2] <- 0
df$hospitalization_24_months[df$hospitalization_24_months == 2] <- 0
df$chronic_disease_24[df$chronic_disease_24 == 2] <- 0
df$type_of_pregnancy[df$type_of_pregnancy == 2] <- 0

#Changing NA to 0 for 
df$number_ABX_treatment_24_months<- replace(df$number_ABX_treatment_24_months, is.na(df$number_ABX_treatment_24_months), 0)
df$hospitalization_24_months<- replace(df$hospitalization_24_months, is.na(df$hospitalization_24_months), 0)
df$prevention_therapy<- replace(df$prevention_therapy, is.na(df$prevention_therapy), 0)
df$number_inhalation<- replace(df$number_inhalation, is.na(df$number_inhalation), 0)
df$AD_treatment_24_month<- replace(df$AD_treatment_24_month, is.na(df$AD_treatment_24_month), 0)
df$steroids_treatment24<- replace(df$steroids_treatment24, is.na(df$steroids_treatment24), 0)
df$number_steroid_treat<- replace(df$number_steroid_treat
, is.na(df$number_steroid_treat), 0)
df$number_of_ABX_first_year<- replace(df$number_of_ABX_first_year, is.na(df$number_of_ABX_first_year), 0)
df <- df %>% drop_na(AD_first_year, FA_first_year, ABX_fisrt_year,food_allergy_24_month, inhalation_24_month, AD_24_month) 

df <- df %>% mutate(SES_group = ifelse(SES>8, '9-10', ifelse(SES>5, '6-8', '1-5')))# create a category SES variable :

#Creating dependent variables with the correct name 
df$FA_second_year <-  df$food_allergy_24_month
df$AD_second_year <- df$AD_24_month
df$HRAD_second_year <- df$inhalation_24_month
df$ABX_second_year <- df$ABX_treatment_24_months
df$feeding_type <- df$study_group_COMEET

#option A of explanatory var:
df <- df %>% mutate(is_atopic_first_year = ifelse((AD_first_year + FA_first_year + ABX_fisrt_year + HRAD_first_year)>0,1,0), 
                    is_atopic_second_year = ifelse((AD_24_month + FA_second_year + ABX_second_year + HRAD_second_year)>0,1,0))

#Choosing explanatory variables: Ones that have a sufficient number of observation before and after Corona
#The distribution of observations according to the timeline of the corona. Before the corona = 1, the first six months of the corona = 2  and after = 3:
group1 <- as.Date("2019-03-01")
group2 <- as.Date("2020-03-11")
group3 <- as.Date("2021-02-07")
df <- df %>% mutate(covid_time = ifelse(DOB < group1, 1, ifelse(DOB < group2, 2, ifelse(DOB<group3, 3, 4))))
df <- df %>% mutate(Born_at = ifelse(covid_time == 1, 'Up to March19', ifelse(covid_time == 2, 'March19 - March20', 'March20 - Feb21')))

df <- df %>% mutate(
        family_asthma = ifelse(mother_asthma + father_asthma + sibling1_asthma + sibling2_asthma + sibling3_asthma + sibling4_asthma >0,1,0),
        family_AD = ifelse(mother_atopic_dermatitis + father_atopic_dermatitis + sibling1_atopic_dermatitis + sibling2_atopic_dermatitis + sibling3_atopic_dermatitis + sibling4_atopic_dermatitis >0,1,0),
        family_Allergic_rhinitis = ifelse(mother_Allergic_rhinitis + father_Allergic_rhinitis + sibling1_Allergic_rhinitis + sibling2_Allergic_rhinitis + sibling3_Allergic_rhinitis + sibling4_Allergic_rhinitis >0,1,0),
        family_FA = ifelse(mother_food_allergy + father_food_allergy + sibling1_food_allergy + sibling2_food_allergy + sibling3_food_allergy + sibling4_food_allergy >0,1,0))
df$DOB_count = as.numeric(as.Date(df$DOB)) - min(as.numeric(as.Date(df$DOB)))
 
# Change categorical variable to factors 
df$feeding_type <- as.factor(df$feeding_type)
df$mother_education <- as.factor(df$mother_education)
df$type_of_pregnancy <- as.factor(df$type_of_pregnancy)
df$nursing_setup_at_12mo <- as.factor(df$nursing_setup_at_12mo)
df$covid_time <- as.factor(df$covid_time)


df <- df %>% drop_na(is_atopic_first_year, is_atopic_second_year)
attach(df)
```

## Sibilings 
```{r}
# insert NA to siblings atopic disease when there is no siblings
df_sib = df %>% mutate(sibling_Atopic_diseases_sum = ifelse(number_of_children ==0, NA, sibling_Atopic_diseases_sum)) 

# Check diff between babies with / without siblings 
sibilings_df <- df_sib %>% select(record_id, number_of_children, is_atopic_first_year, is_atopic_second_year, sibling_Atopic_diseases_sum) %>% 
  mutate(have_siblings = ifelse(number_of_children >0, 1, 0), sibling_Atopic_diseases = ifelse(sibling_Atopic_diseases_sum>0,1,0)) 

sib_group <- sibilings_df %>% group_by(have_siblings) %>%  summarise(n = n(), is_atopic_first_year = sum(is_atopic_first_year), is_atopic_second_year = sum(is_atopic_second_year, na.rm=T))

sib_group$is_atopic_first_year_perc = sib_group$is_atopic_first_year / sib_group$n
sib_group$is_atopic_second_year_perc = sib_group$is_atopic_second_year / sib_group$n
sib_group

# check diff of sick siblings between is_atopic
is_atopic_df1 <- sibilings_df %>% group_by(is_atopic_first_year) %>%  summarise(n = n(), sibling_Atopic_diseases = sum(sibling_Atopic_diseases, na.rm=T), .groups='drop')
is_atopic_df1$perc = is_atopic_df1$sibling_Atopic_diseases/ is_atopic_df1$n
is_atopic_df1


is_atopic_df2 <- sibilings_df %>% group_by(is_atopic_second_year) %>%  summarise(n = n(), sibling_Atopic_diseases = sum(sibling_Atopic_diseases, na.rm=T), .groups='drop')
is_atopic_df2$perc = is_atopic_df2$sibling_Atopic_diseases/ is_atopic_df2$n
is_atopic_df2

# Find correlation between atopic disease to sick siblings
cor(sibilings_df$sibling_Atopic_diseases, sibilings_df$is_atopic_first_year, use = "pairwise.complete.obs")
cor(sibilings_df$sibling_Atopic_diseases, sibilings_df$is_atopic_second_year, use = "pairwise.complete.obs")


### conclusion - it seems like the connection between the two variables is very week, 
### in addition the results at the models below with or without siblings are similar, in my opinion (Shahar) we can skip this variable. 
```


# Descriptive statistics :
```{r}
# SHAHAR
n_obs = length(record_id)
n_atopic1 = sum(is_atopic_first_year)
n_atopic2 = sum(is_atopic_second_year, na.rm = T)

exp_var1 <- df %>% select(c(SES_group, gender, mother_Atopic_diseases, father_Atopic_diseases, sibling_Atopic_diseases,
             mother_education, number_of_children, type_of_pregnancy, covid_time))
exp_var2 <- df %>% select(c(SES, gender, vaccine_12mo, feeding_type, mother_Atopic_diseases, father_Atopic_diseases, sibling_Atopic_diseases, 
             mother_education, number_of_children, type_of_pregnancy, nursing_setup_at_12mo, covid_time))

tablelist1 <- mapply(table, exp_var1)
tablelist2 <- mapply(table, exp_var2)

covid_time_prop <- round(tablelist1[['covid_time']] / sum(tablelist1[['covid_time']]),2)

# boxplot(number_of_children, is_atopic_first_year)

# # EDEN 
# table(covid_time)
# 
# covid_time_prop
# 
# table(covid_time,is_atopic_first_year)
# 
# table(covid_time,is_atopic_second_year)
# 
# summary(number_of_children)
# 
# table(number_of_children, sibling_Atopic_diseases)
```

מספר תצפיות = 1579
מספר חולים בשנה הראשונה = 1035
מספר חולים בשנה השניה = 950

התפךכות בתצפיות בין הקבוצות 
   1    2    3    4 
0.14 0.49 0.35 0.02








## arrange all the colnames by type for more calculation below 
```{r}
# MASBIR
full_binari_exp <- c('Jew/Arab', 'gender', 'Allergy_total', 
                'mother_asthma', 'mother_atopic_dermatitis', 'mother_Allergic_rhinitis', 'mother_food_allergy', 
                'father_asthma', 'father_atopic_dermatitis', 'father_Allergic_rhinitis', 'father_food_allergy', 
                'sibling1_asthma', 'sibling1_atopic_dermatitis', 'sibling1_Allergic_rhinitis', 'sibling1_food_allergy', 
                'sibling2_asthma', 'sibling2_atopic_dermatitis', 'sibling2_Allergic_rhinitis', 'sibling2_food_allergy', 
                'sibling3_asthma', 'sibling3_atopic_dermatitis', 'sibling3_Allergic_rhinitis', 'sibling3_food_allergy', 
                'sibling4_asthma', 'sibling4_atopic_dermatitis', 'sibling4_Allergic_rhinitis', 'sibling4_food_allergy', 
                'vaccine_12mo', 'prevention_therapy', 'steroids_treatment24',
                'hospitalization_24_months', 'chronic_disease_24',
                'family_Atopic_diseases', 'mother_Atopic_diseases', 'father_Atopic_diseases', 'sibling_Atopic_diseases')

binari_exp <- c('Jew/Arab', 'gender', 'Allergy_total','vaccine_12mo', 'prevention_therapy',
                'family_Atopic_diseases', 'mother_Atopic_diseases', 'father_Atopic_diseases', 'sibling_Atopic_diseases')

continues_exp <- c('mother_year_birth', 'DOB', 'mother_education')

discrite_exp <- c('SES', 'number_of_children', 'preganancy_number', 'number_of_ABX_first_year', 'number_inhalation',
                  'family_Atopic_diseases_sum', 'mother_Atopic_diseases_sum', 'father_Atopic_diseases_sum', 'sibling_Atopic_diseases_sum',
                  'sibling_asthma_sum', 'sibling_AD_sum', 'sibling_Allergic_rhinitis_sum', 'sibling_FA_sum')

categoric_exp <- c('type_of_pregnancy', 'nursing_setup_at_12mo', 'covid_time')

#MOOSBAR
exploreatory_variables <- c('HRAD_first_year', 'AD_first_year', 'FA_first_year', 'ABX_fisrt_year',
                            'HRAD_second_year', 'AD_second_year', 'FA_second_year', 'ABX_second_year',
                            'is_atopic_first_year', 'is_atopic_second_year') 


g1 <- as.numeric(as.Date("2019-03-01")) - min(as.numeric(as.Date(df$DOB)))
g2 <- as.numeric(as.Date("2020-03-11")) - min(as.numeric(as.Date(df$DOB)))
g3 <- as.numeric(as.Date("2021-02-07")) - min(as.numeric(as.Date(df$DOB)))

arranged_df <- df %>% arrange(DOB) %>% select(DOB,DOB_count,covid_time)
```

## Two dim table - covid time VS each column 
```{r}
covid_time_table <- function(kpi){
  return(list(perc = round(ftable(covid_time, df[[kpi]]) / rowSums(ftable(covid_time, df[[kpi]])),3),
              n = table(covid_time, df[[kpi]])))}

sapply(binari_exp, covid_time_table, simplify =FALSE)
sapply(discrite_exp, covid_time_table, simplify =FALSE)
sapply(categoric_exp, covid_time_table, simplify =FALSE)
sapply(exploreatory_variables, covid_time_table, simplify =FALSE)
```


## Chi2 test - checking for AB test - results = Failed ! 
```{r, warning=F}
alpha = 0.01
chisq_goodness <- function(kpi){
  t <- chisq.test(df[[kpi]], df$covid_time)
  # cat(kpi, '| p_val' , t$p.value, '| SRM =' , t$p.value < alpha, '\n ')
  return(list('p_val'=t$p.value, 'SRM'= t$p.value < alpha))} 


# sapply(binari_exp, chisq_goodness, simplify =FALSE)
# sapply(discrite_exp, chisq_goodness, simplify =FALSE)
# sapply(categoric_exp, chisq_goodness, simplify =FALSE)
# sapply(exploreatory_variables, chisq_goodness, simplify =FALSE)
```



# Regression 

## Logistic regression
```{r}
df_mdl <- df %>% drop_na(SES_group, gender, type_of_pregnancy, mother_education, vaccine_12mo, nursing_setup_at_12mo, 
                         is_atopic_first_year, is_atopic_second_year)

mdl1 <- glm(is_atopic_first_year ~ SES_group + gender + type_of_pregnancy +
               mother_Atopic_diseases + father_Atopic_diseases +
               mother_year_birth + mother_education + number_of_children + covid_time  , data = df_mdl, family = "binomial")

mdl2 <- glm(is_atopic_second_year ~ SES_group + gender + 
             mother_Atopic_diseases + father_Atopic_diseases +
             mother_year_birth + mother_education + number_of_children + 
             type_of_pregnancy  + vaccine_12mo  + nursing_setup_at_12mo + covid_time, data=df_mdl, family='binomial')

df_mdl$mdl1 <- predict(mdl1)
df_mdl$mdl2 <- predict(mdl2)

df_group <- df_mdl %>% group_by(covid_time) %>% summarise(mdl1_avg = mean(mdl1, na.rm=T),  
                                                          mdl2_avg = mean(mdl2, na.rm=T))
df_mdl <- left_join(df_mdl, df_group, by = 'covid_time')

summary(mdl1)
```


## GAM : 
```{r}
gam1 <- gam(is_atopic_first_year ~ SES_group + gender + type_of_pregnancy +
               mother_Atopic_diseases + father_Atopic_diseases +
               mother_year_birth + mother_education + number_of_children +
               s(DOB_count, bs='cr') ,data=df_mdl, family=binomial)

gam2 <- gam(is_atopic_second_year ~ SES_group + gender + 
             mother_Atopic_diseases + father_Atopic_diseases +
             mother_year_birth + mother_education + number_of_children + 
             type_of_pregnancy  + vaccine_12mo  + nursing_setup_at_12mo + s(DOB_count, bs='cr') ,data=df_mdl, family=binomial)

df_mdl$gam1 <- predict.gam(gam1)
df_mdl$gam2 <- predict.gam(gam2)

summary.gam(gam1)
```

# Graphs

```{r}
# graph of the prob for is atopic 1 year
p_prob <- plot(getViz(gam1), shift = coef(gam1)[1], trans = plogis, pages = 1)
p_prob$plots[[1]]$data$fit
```


```{r}
library(voxel)
library(lmerTest)
library(mgcViz)

p <- plot(getViz(gam1))
x <- p$plots[[1]]$data$fit$x
y <- p$plots[[1]]$data$fit$y

gam_mdl_df <- p$plots[[1]]$data$fit
gam_mdl_df <- gam_mdl_df %>% mutate(Born_at = if_else(x< g1, 'Up to March19', ifelse(x<g2, 'March19 - March20', ifelse(x<g3,'March20 - Feb21','From Feb21'))))
gam_mdl_df$Born_at <- as.factor(gam_mdl_df$Born_at)

gam_mdl_df <- gam_mdl_df %>% 
  mutate(logistic_reg = if_else(x < g1, 0, ifelse(x<g2, mdl1$coefficients['covid_time2'] , ifelse(x<g3,mdl1$coefficients['covid_time3'], mdl1$coefficients['covid_time4']))))

plus = min(as.numeric(as.Date(df$DOB)))
gam_mdl_df <- gam_mdl_df %>% mutate(Date = as.Date(x + plus,  origin = '1970-01-01'))


ggplot(gam_mdl_df, aes(Date,y)) + geom_line(aes(col = Born_at), size = 1) +
  geom_line(aes(Date, logistic_reg, col = Born_at), size = 1) + 
  geom_ribbon(aes(ymin = y - se*qnorm(0.975), ymax = y + se*qnorm(0.975), y=NULL), alpha = 0.1 ) +
  theme_bw() + 
  labs(title = 'GAM model for atopic in first year', x = 'Date of Birth', y = 'is_atopic_first_year') +
  scale_color_manual(values=viridis(4))
```

## Create Graphs - Old
```{r, error=F, warning=F, message=F}
# design = theme_bw() 

# First year - No Sibilings
# ggplot(df_mdl, aes(DOB, gam1)) + geom_smooth(aes(col = covid_time)) + geom_line(aes(DOB, mdl1_avg, col = covid_time)) + 
#   design + labs(title = 'Atopic disease risk for babies up to one year old by DOB', x = 'Date of Birth', y = 'Atopic disease probability') +
#   scale_color_manual(values=viridis(4), name = "Born at", labels =c('Up to March19','March19 - March20','March20 - Feb21', 'From Feb21' ))
# 
# # Seconde year - No Sibilings
# ggplot(df_mdl, aes(DOB, gam2)) + 
#   geom_smooth(aes(col = covid_time)) + 
#   geom_line(aes(DOB, mdl2_avg, col = covid_time)) +
#   design + 
#   labs(title = 'Atopic disease risk for 1-2 year old babies by DOB', x = 'Date of Birth', y = 'Atopic disease probability') +
#   scale_color_manual(values=viridis(4), name = "Born at", labels =c('Up to March19','March19 - March20','March20 - Feb21', 'From Feb21' ))+ 
#   ylim(0.5,1)
```



# Predictions:

## Typical person :

```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]}


n <- length(df_mdl$record_id)

typical_person <- data.frame(SES_group = rep(getmode(SES_group),n), 
                             gender = rep(getmode(gender),n),
                             type_of_pregnancy = rep(getmode(type_of_pregnancy),n),
                             mother_Atopic_diseases = rep(getmode(mother_Atopic_diseases),n),
                             father_Atopic_diseases = rep(getmode(father_Atopic_diseases),n),
                             mother_year_birth = rep(mean(mother_year_birth),n),
                             mother_education = rep(getmode(mother_education),n),
                             number_of_children = rep(getmode(number_of_children),n))


typical_person <- typical_person %>% mutate(DOB = df_mdl$DOB, DOB_count = df_mdl$DOB_count)
typical_person$predict <- predict.gam(gam1, typical_person, type = 'response')


ggplot(typical_person, aes(DOB, predict)) + geom_line()
```

