# Import
```{r, echo=F, message=F}
# libraries:
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape)
library(glue)
options(scipen=5, digits = 5)
```

# Prepare the data to analysis 
```{r}
# Reading the file:
# dir <- "C:/Users/Merav/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_3.1.23.xlsx"
dir <- "C:/Users/edena/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_3.1.23.xlsx"
#dir<-"C:/Users/Shahar/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_3.1.23.xlsx"
    
df <- read_excel(dir, sheet = 1)
df <- df %>% mutate(DOB = as.Date(DOB)) 

# Arranging df:
# Filtering only participants that completed the survey:
df <- df %>% filter(is.na(lost_of_followup))
# Filtering only participants that completed the second year:
df <- df %>% filter(month_survey_complete == 2)

# Removing non relevant columns: from type_of_allergy to 12_month_FA_type_allergy
#remove_columns = colnames(df)[51:88]
df <- df[,-c(51:88)]

# Removing columns with allot on NA:
not_relevant <- c('allergist_followup', 'last_spt', 'epipen', 'mistake_exposure', 'reaction_type24', 'diagnosis_by', 'lost_of_followup',"mother_no_comorbidities","mother_other","father_no_comorbidities","father_other","sibling1_other","sibling1_no_comorbidities","sibling2_other","sibling2_no_comorbidities","sibling3_other","sibling3_no_comorbidities","sibling4_other","sibling4_no_comorbidities")
df <- df[-which(colnames(df) %in% not_relevant)]

# Adding binary columns 
# A. Did any one from family (mother, father, sib) have *any* of the disease?
df$family_Atopic_diseases_sum <- rowSums(df[,14:37])

# B. Did somebody from family have a specific disease?
# NEED TO FILL IN

# C. Did a disease appear in a specific family member?
df$mother_Atopic_diseases_sum <- rowSums(df[,14:17])
df$father_Atopic_diseases_sum <- rowSums(df[,18:21])
df$sibling_Atopic_diseases_sum <- rowSums(df[,22:37])

# Sibilings
df <- df %>% mutate(sibling_asthma_sum = sibling1_asthma + sibling2_asthma + sibling3_asthma + sibling4_asthma,
                    sibling_AD_sum = sibling1_atopic_dermatitis + sibling2_atopic_dermatitis + sibling3_atopic_dermatitis + sibling4_atopic_dermatitis,
                    sibling_Allergic_rhinitis_sum = sibling1_Allergic_rhinitis + sibling2_Allergic_rhinitis + sibling3_Allergic_rhinitis + sibling4_Allergic_rhinitis,
                    sibling_FA_sum = sibling1_food_allergy + sibling2_food_allergy + sibling3_food_allergy + sibling4_food_allergy)
              
df <- df %>% mutate(family_Atopic_diseases = ifelse(family_Atopic_diseases_sum >0,1,0),
                    mother_Atopic_diseases = ifelse(mother_Atopic_diseases_sum>0,1,0),
                    father_Atopic_diseases = ifelse(father_Atopic_diseases_sum > 0, 1,0),
                    sibling_Atopic_diseases = ifelse(sibling_Atopic_diseases_sum > 0 ,1, 0))

# Changing to binary variables:
df$gender[df$gender == 2] <- 0
df$`Jew/Arab`[df$`Jew/Arab` == 2] <- 0
df$food_allergy_24_month[df$food_allergy_24_month == 2] <- 0
df$inhalation_24_month[df$inhalation_24_month == 2] <- 0
df$prevention_therapy[df$prevention_therapy == 2] <- 0
df$AD_24_month[df$AD_24_month == 2] <- 0
df$steroids_treatment24[df$steroids_treatment24 == 2] <- 0
df$ABX_treatment_24_months[df$ABX_treatment_24_months == 2] <- 0
df$hospitalization_24_months[df$hospitalization_24_months == 2] <- 0
df$chronic_disease_24[df$chronic_disease_24 == 2] <- 0
df$type_of_pregnancy[df$type_of_pregnancy == 2] <- 0

#Changing NA to 0 for 
df$number_ABX_treatment_24_months<- replace(df$number_ABX_treatment_24_months, is.na(df$number_ABX_treatment_24_months), 0)
df$hospitalization_24_months<- replace(df$hospitalization_24_months, is.na(df$hospitalization_24_months), 0)

df$prevention_therapy<- replace(df$prevention_therapy, is.na(df$prevention_therapy), 0)
df$number_inhalation<- replace(df$number_inhalation, is.na(df$number_inhalation), 0)

df$AD_treatment_24_month<- replace(df$AD_treatment_24_month, is.na(df$AD_treatment_24_month), 0)
df$steroids_treatment24<- replace(df$steroids_treatment24, is.na(df$steroids_treatment24), 0)
df$number_steroid_treat<- replace(df$number_steroid_treat
, is.na(df$number_steroid_treat), 0)

df$number_of_ABX_first_year<- replace(df$number_of_ABX_first_year, is.na(df$number_of_ABX_first_year), 0)

df <- df %>% drop_na(AD_first_year, FA_first_year, ABX_fisrt_year,food_allergy_24_month, inhalation_24_month, AD_24_month) 

#Creating dependent variables:
df$FA_second_year <-  df$food_allergy_24_month
df$AD_second_year <- df$AD_24_month
df$HRAD_second_year <- df$inhalation_24_month
df$ABX_second_year <- df$ABX_treatment_24_months

#option A of explanatory var:
df <- df %>% mutate(is_atopic_first_year = ifelse((AD_first_year + FA_first_year + ABX_fisrt_year + HRAD_first_year)>0,1,0), 
                    is_atopic_second_year = ifelse((AD_24_month + FA_second_year + ABX_second_year + HRAD_second_year)>0,1,0))

#option B of explanatory var:
#From mail

#Choosing explanatory variables: Ones that have a sufficient number of observation before and after Corona
#The distribution of observations according to the timeline of the corona. Before the corona = 1, the first six months of the corona = 2  and after = 3:

group1 <- as.Date("2019-03-01")
group2 <- as.Date("2020-03-11")
group3 <- as.Date("2021-02-07")
df <- df %>% mutate(covid_time = ifelse(DOB < group1, 1, ifelse(DOB < group2, 2, ifelse(DOB<group3, 3, 4))))
table(df$covid_time)
df <- df %>% mutate(Born_at = ifelse(covid_time == 1, 'Up to March19', ifelse(covid_time == 2, 'March19 - March20', 'March20 - Feb21')))

df <- df %>% mutate(
        family_asthma = ifelse(mother_asthma + father_asthma + sibling1_asthma + sibling2_asthma + sibling3_asthma + sibling4_asthma >0,1,0),
        family_AD = ifelse(mother_atopic_dermatitis + father_atopic_dermatitis + sibling1_atopic_dermatitis + sibling2_atopic_dermatitis + sibling3_atopic_dermatitis + sibling4_atopic_dermatitis >0,1,0),
        family_Allergic_rhinitis = ifelse(mother_Allergic_rhinitis + father_Allergic_rhinitis + sibling1_Allergic_rhinitis + sibling2_Allergic_rhinitis + sibling3_Allergic_rhinitis + sibling4_Allergic_rhinitis >0,1,0),
        family_FA = ifelse(mother_food_allergy + father_food_allergy + sibling1_food_allergy + sibling2_food_allergy + sibling3_food_allergy + sibling4_food_allergy >0,1,0))


# at this moment there is no sutible data from group 4 - post corona 
attach(df)
```

# Summary of the data:
```{r}
# Frequency table for each column
tablelist <- mapply(table, df %>% select(-c(record_id, DOB, mother_year_birth)))
# tablelist
```


# NA count 
```{r}
library(pander)
# How many NA are in each column
NA_count <- as.data.frame(t(df %>% summarise_all(funs(sum(is.na(.)))))) %>% dplyr::rename("amount of NA" = 1) 
NA_count <- NA_count %>% mutate(Name = rownames(NA_count))
NA_column = NA_count %>% filter(`amount of NA` >0) %>% arrange(`amount of NA`)

n = 1439
xxx <- df[is.na(df$mother_education),]

round(ftable(xxx$covid_time, xxx$is_atopic_first_year) / rowSums(ftable(xxx$covid_time, xxx$is_atopic_first_year)),3)
# ftable(xxx$covid_time, xxx$is_atopic_first_year)/ #166
ftable(xxx$covid_time, xxx$is_atopic_first_year)

round(ftable(covid_time, is_atopic_first_year) / rowSums(ftable(covid_time, is_atopic_first_year)),3)
# ftable(covid_time, is_atopic_first_year) /n
ftable(covid_time, is_atopic_first_year)

```


# arrange all the colnames by type
```{r}
# MASBIR
full_binari_exp <- c('Jew/Arab', 'gender', 'Allergy_total', 
                'mother_asthma', 'mother_atopic_dermatitis', 'mother_Allergic_rhinitis', 'mother_food_allergy', 
                'father_asthma', 'father_atopic_dermatitis', 'father_Allergic_rhinitis', 'father_food_allergy', 
                'sibling1_asthma', 'sibling1_atopic_dermatitis', 'sibling1_Allergic_rhinitis', 'sibling1_food_allergy', 
                'sibling2_asthma', 'sibling2_atopic_dermatitis', 'sibling2_Allergic_rhinitis', 'sibling2_food_allergy', 
                'sibling3_asthma', 'sibling3_atopic_dermatitis', 'sibling3_Allergic_rhinitis', 'sibling3_food_allergy', 
                'sibling4_asthma', 'sibling4_atopic_dermatitis', 'sibling4_Allergic_rhinitis', 'sibling4_food_allergy', 
                'vaccine_12mo', 'prevention_therapy', 'steroids_treatment24',
                'hospitalization_24_months', 'chronic_disease_24',
                'family_Atopic_diseases', 'mother_Atopic_diseases', 'father_Atopic_diseases', 'sibling_Atopic_diseases')

binari_exp <- c('Jew/Arab', 'gender', 'Allergy_total','vaccine_12mo', 'prevention_therapy',
                'family_Atopic_diseases', 'mother_Atopic_diseases', 'father_Atopic_diseases', 'sibling_Atopic_diseases')

continues_exp <- c('mother_year_birth', 'DOB', 'mother_education')

discrite_exp <- c('SES', 'number_of_children', 'preganancy_number', 'number_of_ABX_first_year', 'number_inhalation',
                  'family_Atopic_diseases_sum', 'mother_Atopic_diseases_sum', 'father_Atopic_diseases_sum', 'sibling_Atopic_diseases_sum',
                  'sibling_asthma_sum', 'sibling_AD_sum', 'sibling_Allergic_rhinitis_sum', 'sibling_FA_sum')

categoric_exp <- c('type_of_pregnancy', 'nursing_setup_at_12mo', 'covid_time')

#MOOSBAR
exploreatory_variables <- c('HRAD_first_year', 'AD_first_year', 'FA_first_year', 'ABX_fisrt_year',
                            'HRAD_second_year', 'AD_second_year', 'FA_second_year', 'ABX_second_year',
                            'is_atopic_first_year', 'is_atopic_second_year') 
```


# Two dim table - covid time VS each column 
```{r}
# table(is_atopic_first_year, covid_time)
# table(is_atopic_second_year, covid_time)

covid_time_table <- function(kpi){
  return(list(perc = round(ftable(covid_time, df[[kpi]]) / rowSums(ftable(covid_time, df[[kpi]])),3),
              n = table(covid_time, df[[kpi]])))}

# sapply(binari_exp, covid_time_table, simplify =FALSE)
# sapply(discrite_exp, covid_time_table, simplify =FALSE)
# sapply(categoric_exp, covid_time_table, simplify =FALSE)
sapply(exploreatory_variables, covid_time_table, simplify =FALSE)
```


```{r}
variable_list = c('is_atopic_first_year', 'is_atopic_second_year')

covid_time_graphs <- function(kpi){
  return (ggplot(df, aes(x=covid_time, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_col() + ggtitle(kpi))}
lapply(variable_list, covid_time_graphs)
```




# Graphs and tables that have been checked and relevant :
## covid time VS no_is_atopic
```{r, warning=F}
no_is_atopic = c('HRAD_first_year', 'AD_first_year', 'FA_first_year', 'ABX_fisrt_year',
                            'HRAD_second_year', 'AD_second_year', 'FA_second_year', 'ABX_second_year')

atopic_longer_count <- df %>% select(c(no_is_atopic, covid_time,Born_at)) %>% drop_na() %>% group_by(covid_time, Born_at) %>% summarise_if(is.numeric, mean) %>% pivot_longer(cols  =no_is_atopic, names_to = "chronic_disease", values_to = "mean") 

ggplot(atopic_longer_count, aes(covid_time, y = mean, fill = as.factor(covid_time))) + 
  geom_col() + facet_wrap(~chronic_disease,4,2) + geom_text(aes(label = round(atopic_longer_count$mean,3)), vjust = 0, size = 3) + theme_bw() +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'),
                    name = "Born at", labels =c('Up to March19','March19 - March20','March20 - Feb21')) + 
  xlab("") + theme(axis.text.x = element_text( color = 'white')) + ylim(c(0,0.7))
```
## covid time VS is_atopic

```{r}
is_atopic = c('is_atopic_first_year', 'is_atopic_second_year')

is_atopic_longer <- df %>% select(c(is_atopic, covid_time,Born_at)) %>% drop_na() %>% group_by(covid_time, Born_at) %>% summarise_if(is.numeric, mean) %>% pivot_longer(cols = is_atopic, names_to = "chronic_disease", values_to = "mean") 

ggplot(is_atopic_longer, aes(covid_time, y = mean, fill = as.factor(covid_time))) + 
  geom_col() + theme_bw() + facet_wrap(~chronic_disease) + 
  scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'),
                    name = "Born at", labels =c('Up to March19','March19 - March20','March20 - Feb21')) + 
  xlab("")+  ylab("proportion") + theme(axis.text.x = element_text(color = 'white')) +
  geom_text(aes(label = round(is_atopic_longer$mean,3)), vjust = 0, size = 3)
```
## covid time VS family 

```{r, warning=F}
atopic_binari_exp <- c('family_asthma', 'family_AD', 'family_Allergic_rhinitis', 'family_FA')
new_df_longer <- df %>% select(c(atopic_binari_exp, covid_time, Born_at)) %>% drop_na() %>% group_by(covid_time, Born_at) %>% summarise_if(is.numeric, mean) %>% pivot_longer(cols = atopic_binari_exp, names_to = "chronic_disease", values_to = "mean") 

ggplot(new_df_longer, aes(covid_time, y = mean, fill = as.factor(covid_time))) + 
  geom_col() + theme_bw() +
  facet_wrap(~chronic_disease) + 
  scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'),
                    name = "Born at", labels =c('Up to March19','March19 - March20','March20 - Feb21')) + 
  xlab("") + ylab("proportion") + theme(axis.text.x = element_text(color = 'white')) +
  geom_text(aes(label = round(new_df_longer$mean,3)), vjust = 0, size = 3) + ylim(c(0, 0.25))
```

## covid time VS number_of_treatments
צריך להפוך את זה לגרף מוערם 

צריך לשים לב שאני פילטרתי את מספר בטיפולים = 0 כי אני רוצה לראות את היחס בין מי שכן קיבל טיפול, האם החולי הפך לאגרסיבי יותר ? 
 האם כדאי להראות את זה ביחס?
```{r}
# how many treatment was given in different covid time groups 
treatments_kpi <- c('number_of_ABX_first_year', 'number_inhalation','number_steroid_treat')

# proportion tables 
# treatments_prop <- function(kpi){
#   data <- df %>% select(covid_time, kpi) %>% filter(df[[kpi]] > 0)
#   return(round(ftable(data[['covid_time']], data[[kpi]]) / rowSums(ftable(data[['covid_time']], data[[kpi]])),3))}
# sapply(treatments_kpi, treatments_prop, simplify = FALSE)



ggplot(df %>% filter(number_of_ABX_first_year >0 )%>% mutate(covid_time = as.factor(covid_time)), aes(x=covid_time, fill=as.factor(number_of_ABX_first_year))) + geom_bar(position = "fill") + 
  labs( title = 'covid time VS number_of_ABX_first_year', x = 'covid time group', y = 'count', fill = 'number of antib...')+ scale_x_discrete(labels = c('Up to March19','March19 - March20','March20 - Feb21')) + theme_bw()
df %>% filter(number_of_ABX_first_year >0 )%>% group_by(covid_time) %>% summarise(avg_of_ABX_first_year = mean(number_of_ABX_first_year))


ggplot(df %>% filter(number_inhalation >0 )%>% mutate(covid_time = as.factor(covid_time)), aes(x=covid_time, fill=as.factor(number_inhalation))) + geom_bar(position = "fill") + 
  labs( title = 'covid time VS number_inhalation', x = 'covid time group', y = 'count', fill = 'number of antib...')+ scale_x_discrete(labels = c('Up to March19','March19 - March20','March20 - Feb21')) + theme_bw()
df %>% filter(number_inhalation >0 )%>% group_by(covid_time) %>% summarise(avg_number_inhalation = mean(number_inhalation))


ggplot(df %>% filter(number_steroid_treat >0 ) %>% mutate(covid_time = as.factor(covid_time)), aes(x=covid_time, fill=as.factor(number_steroid_treat))) + geom_bar(position = "fill") + 
  labs( title = 'covid time VS steroid_treat', x = 'covid time group', y = 'count', fill = 'number of antib...')+ scale_x_discrete(labels = c('Up to March19','March19 - March20','March20 - Feb21')) + theme_bw() 
df %>% filter(number_steroid_treat >0 )%>% group_by(covid_time) %>% summarise(avg_steroid_treat = mean(number_steroid_treat))
```



## גרף של גיל האמא 
```{r}
ggplot(df, aes(mother_year_birth, fill = as.factor(covid_time))) + geom_density(alpha = 0.3) + theme_bw() + labs(title = 'mother year of birth VS. covid time', fill = 'covid time group')
```

##covid time vs ther rest of categorical variables:
```{r}
#SES
p = as.data.frame(with(df,table(covid_time,SES))/rowSums(with(df,table(covid_time,SES))))

ggplot(p,aes(x=as.factor(covid_time), y= Freq, fill = as.factor(SES))) +
    geom_col() + scale_fill_discrete(name = "SES")+ xlab("covid_time") + ylab("percentage")
#number_of_children
p = as.data.frame(with(df,table(covid_time,number_of_children))/rowSums(with(df,table(covid_time,number_of_children))))

ggplot(p,aes(x=as.factor(covid_time), y= Freq, fill = as.factor(number_of_children))) +
    geom_col() + scale_fill_discrete(name = "number_of_children")+ xlab("covid_time") + ylab("percentage")
#preganancy_number
p = as.data.frame(with(df,table(covid_time,preganancy_number))/rowSums(with(df,table(covid_time,preganancy_number))))

ggplot(p,aes(x=as.factor(covid_time), y= Freq, fill = as.factor(preganancy_number))) +
    geom_col() + scale_fill_discrete(name = "preganancy_number")+ xlab("covid_time") + ylab("percentage")
#type_of_pregnancy
p = as.data.frame(with(df,table(covid_time,type_of_pregnancy))/rowSums(with(df,table(covid_time,type_of_pregnancy))))

ggplot(p,aes(x=as.factor(covid_time), y= Freq, fill = as.factor(type_of_pregnancy))) +
    geom_col() + scale_fill_discrete(name = "type_of_pregnancy")+ xlab("covid_time") + ylab("percentage")
#nursing_setup_at_12mo
p = as.data.frame(with(df,table(covid_time,nursing_setup_at_12mo))/rowSums(with(df,table(covid_time,nursing_setup_at_12mo))))

ggplot(p,aes(x=as.factor(covid_time), y= Freq, fill = as.factor(nursing_setup_at_12mo))) +
    geom_col() + scale_fill_discrete(name = "nursing_setup_at_12mo")+ xlab("covid_time") + ylab("percentage")
```

##covid time vs ther rest of binary variables:
עשיתי בלי חיסון שנה ראשונה כי כמעט כולם קיבלו ואין מה להציג.

```{r}
other_var <- c('SES', 'number_of_children', 'preganancy_number','type_of_pregnancy', 'nursing_setup_at_12mo')

binary_longer <- df %>% select(c(prevention_therapy,steroids_treatment24, hospitalization_24_months,chronic_disease_24 , covid_time,Born_at)) %>% drop_na() %>% group_by(covid_time, Born_at) %>% summarise_if(is.numeric, mean) %>% gather(variables, mean, prevention_therapy:chronic_disease_24, factor_key=TRUE)# pivot_longer(cols = is_atopic, names_to = "chronic_disease", values_to = "mean") 

ggplot(binary_longer, aes(covid_time, y = mean, fill = as.factor(covid_time))) + 
  geom_col() + theme_bw() + facet_wrap(~variables) + 
  scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'),
                    name = "Born at", labels =c('Up to March19','March19 - March20','March20 - Feb21')) + 
  xlab("")+  ylab("proportion") + theme(axis.text.x = element_text(color = 'white')) +
  geom_text(aes(label = round(binary_longer$mean,3)), vjust = 0, size = 3)

```

## השכלה של האמא 
```{r}
ggplot(df, aes(covid_time, fill = as.factor(mother_education))) + geom_bar(position = "fill")
```


## כל הקטגוריים והבינאריים מול הקבוצות קורונה 

## number_of_children

# Not relevant graphs : 

# Graphs that need to be checked : 

##Age of mother vs is atopic
```{r}
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(is_atopic_first_year), fill=as.factor(is_atopic_first_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(is_atopic_second_year), fill=as.factor(is_atopic_second_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(AD_first_year), fill=as.factor(AD_first_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(AD_24_month), fill=as.factor(AD_24_month))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(FA_first_year), fill=as.factor(FA_first_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(food_allergy_24_month), fill=as.factor(food_allergy_24_month))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(ABX_fisrt_year), fill=as.factor(ABX_fisrt_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(inhalation_24_month), fill=as.factor(inhalation_24_month))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(is_atopic_first_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(is_atopic_second_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(AD_first_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(AD_24_month), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(FA_first_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(food_allergy_24_month), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(ABX_fisrt_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(inhalation_24_month), y=mother_age)) + 
#   geom_boxplot()
# 
```


##mother_education va is atopic
```{r,fig.width=10}
# plot_df_4 <- df %>% select(mother_education,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(mother_education) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_4,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(mother_education))) +
#       geom_col(position = "dodge") + scale_fill_discrete(name = "mother_education")+ xlab("mother_education") + ylab("percent diseas") #+scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))
```

##number_of_children vs is atopic
```{r,fig.width=10}
# plot_df_5 <- df %>% select(number_of_children,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(number_of_children) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_5,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(number_of_children))) +
#     geom_col(position = "dodge") + scale_fill_discrete(name = "number_of_children")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))
```

##preganancy_number vs is atopic
```{r,fig.width=10}
# plot_df_6 <- df %>% select(preganancy_number,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(preganancy_number) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_6,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(preganancy_number))) +
#     geom_col(position = "dodge") + scale_fill_discrete(name = "preganancy_number")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

```

## type_of_pregnancy vs ia atopic
```{r,fig.width=7}
# plot_df_7 <- df %>% select(type_of_pregnancy,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(type_of_pregnancy) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_7,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(type_of_pregnancy))) +
#     geom_col(position = "dodge") + scale_fill_discrete(name = "type_of_pregnancy")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))
```




##nursing_setup_at_12mo vs is atopic
```{r,fig.width=12}
plot_df_10 <- df %>% select(nursing_setup_at_12mo,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(nursing_setup_at_12mo) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)

ggplot(plot_df_10,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(nursing_setup_at_12mo))) +
    geom_col(position = "dodge") + scale_fill_discrete(name = "nursing_setup_at_12mo")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

```

##binary vs catagoric:
```{r, warning=F}
cat <- c("SES","mother_education", "number_of_children","preganancy_number","type_of_pregnancy","number_of_ABX_first_year","nursing_setup_at_12mo")

binary = c("father_food_allergy","HRAD_first_year","family_Atopic_diseases","mother_Atopic_diseases","father_Atopic_diseases","sibling_Atopic_diseases")


for (a in cat) {
  for (b in binary) {
    a_index <- grep(paste0("^",a,"$") , colnames(df))
    b_index <- grep(paste0("^",b,"$"), colnames(df))
    x <- sort(unique(unlist(df[,a_index])))
    y <- sort(unique(unlist(df[,b_index])))
    data <- expand.grid(X=x, Y=y)
    data$Z <- c(table(unlist(df[,a_index]),unlist(df[,b_index])))
    p = ggplot(data, aes(as.factor(X), as.factor(Y), fill= Z)) +
      geom_tile()+
      scale_fill_gradient(low="red", high="blue") + xlab(a) + ylab(b)
    print(p)
  }
}

gg <- function(kpi){ list(percents = round(ftable(df[[kpi]], family_Atopic_diseases) / rowSums(ftable(df[[kpi]], family_Atopic_diseases)),3), n = ftable(df[[kpi]], family_Atopic_diseases))}
sapply(cat, gg,  simplify =FALSE)
```


##binary vs continues:
```{r, fig.width=6}
for (c in binary) {
  c_index <- grep(paste0("^",c,"$"), colnames(df))
  p =ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(unlist(df[,c_index])), fill=as.factor(unlist(df[,c_index])))) +
    geom_density(adjust=1.5, alpha=.4)+ scale_fill_discrete(name = c)
  print(p)
  
}
```


## continues exploratoty variables - DOB Vs Atopic deasese 

```{r, warning=F}
exloretory_variables = list('AD_first_year', 'AD_24_month', 'FA_first_year', 'food_allergy_24_month', 'ABX_fisrt_year', 'inhalation_24_month', 'is_atopic_first_year', 'is_atopic_second_year')

DOB_explanatory <- function(kpi){
  return (ggplot(df, aes(x=DOB, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_density(adjust=1.5, alpha=.4) + ggtitle(kpi))}
lapply(exloretory_variables, DOB_explanatory)
```


## Continues exploratoty VS Categoty exploratoty
```{r, warning=F}
# Categoric_variables <- c('SES', 'Jew/Arab', 'gender', 'study_group_COMEET', 'mother_education', 'number_of_children', 'preganancy_number', 'type_of_pregnancy', "family_Atopic_diseases", "mother_Atopic_diseases", "father_Atopic_diseases", "sibling_Atopic_diseases" )
# 
# DOB_explanatory_dist <- function(kpi){
#   return (ggplot(df, aes(x=DOB, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_density(adjust=1.5, alpha=.4) + ggtitle(kpi))}
# lapply(Categoric_variables , DOB_explanatory_dist)
# 
# DOB_explanatory_dist <- function(kpi){
#   return (ggplot(df, aes(x=DOB, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_boxplot(adjust=1.5, alpha=.4) + ggtitle(kpi))}
# lapply(Categoric_variables , DOB_explanatory_dist)
```


## Categoty exploratoty VS Categoty exploratoty
```{r, warning=F}
# # categoric_exp
# # discrite_exp
# 
# family <- function(kpi){
#   return (ggplot(df, aes(family_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , family)
# 
# mother <- function(kpi){
#   return (ggplot(df, aes(mother_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , mother)
# 
# father <- function(kpi){
#   return (ggplot(df, aes(father_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , father)
# 
# sibling <- function(kpi){
#   return (ggplot(df, aes(sibling_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , sibling)
```




## Binary VS binary

```{r}
library(corrplot)
df_binary_first<- df %>% select("Jew/Arab", "gender", "type_of_pregnancy", "family_Atopic_diseases", "mother_Atopic_diseases","father_Atopic_diseases", "sibling_Atopic_diseases", "HRAD_first_year", "AD_first_year", "FA_first_year", "ABX_fisrt_year", "is_atopic_first_year")

df_binary_second<- df %>% select("Jew/Arab", "gender", "type_of_pregnancy", "family_Atopic_diseases", "mother_Atopic_diseases","father_Atopic_diseases", "sibling_Atopic_diseases", "vaccine_12mo", "prevention_therapy", "vaccines_24month", "steroids_treatment24", "hospitalization_24_months", "chronic_disease_24", "food_allergy_24_month", "inhalation_24_month", "AD_24_month", "is_atopic_second_year")

cor_1<- cor(df_binary_first, method = c("spearman"), use="pairwise.complete.obs")
corrplot(cor_1, method = 'color', order = 'FPC', type = 'lower', diag = FALSE, col = COL2('BrBG'))

cor_2<- cor(df_binary_second, method = c("spearman"), use="pairwise.complete.obs")
corrplot(cor_2, method = 'color', order = 'FPC', type = 'lower', diag = FALSE, col=COL2('BrBG'))
```









