```{r, echo=F, message=F}
# libraries:
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape)
library(glue)
options(scipen=5, digits = 5)
```

# Descriptive statistics

```{r}
# Reading the file:
# dir <- "C:/Users/Merav/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_3.1.23.xlsx"
# dir <- "C:/Users/edena/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_3.1.23.xlsx"
dir<-"C:/Users/Shahar/Documents/GitHub/Statistics-Project/atopic_comorbidities_first_second_year_3.1.23.xlsx"
    
df <- read_excel(dir, sheet = 1)
df <- df %>% mutate(DOB = as.Date(DOB)) 

# Arranging df:
# Filtering only participants that completed the survey:
df <- df %>% filter(is.na(lost_of_followup))
# Filtering only participants that completed the second year:
df <- df %>% filter(month_survey_complete == 2)

# Removing non relevant columns: from type_of_allergy to 12_month_FA_type_allergy
#remove_columns = colnames(df)[51:88]
df <- df[,-c(51:88)]

# Removing columns with allot on NA:
not_relevant <- c('allergist_followup', 'last_spt', 'epipen', 'mistake_exposure', 'reaction_type24', 'diagnosis_by', 'lost_of_followup',"mother_no_comorbidities","mother_other","father_no_comorbidities","father_other","sibling1_other","sibling1_no_comorbidities","sibling2_other","sibling2_no_comorbidities","sibling3_other","sibling3_no_comorbidities","sibling4_other","sibling4_no_comorbidities")
df <- df[-which(colnames(df) %in% not_relevant)]

# Adding binary columns 
# A. Did any one from family (mother, father, sib) have *any* of the disease?
df$family_Atopic_diseases_sum <- rowSums(df[,14:37])

# B. Did somebody from family have a specific disease?
# NEED TO FILL IN

# C. Did a disease appear in a specific family member?
df$mother_Atopic_diseases_sum <- rowSums(df[,14:17])
df$father_Atopic_diseases_sum <- rowSums(df[,18:21])
df$sibling_Atopic_diseases_sum <- rowSums(df[,22:37])

# Sibilings
df <- df %>% mutate(sibling_asthma_sum = sibling1_asthma + sibling2_asthma + sibling3_asthma + sibling4_asthma,
                    sibling_AD_sum = sibling1_atopic_dermatitis + sibling2_atopic_dermatitis + sibling3_atopic_dermatitis + sibling4_atopic_dermatitis,
                    sibling_Allergic_rhinitis_sum = sibling1_Allergic_rhinitis + sibling2_Allergic_rhinitis + sibling3_Allergic_rhinitis + sibling4_Allergic_rhinitis,
                    sibling_FA_sum = sibling1_food_allergy + sibling2_food_allergy + sibling3_food_allergy + sibling4_food_allergy)
              
df <- df %>% mutate(family_Atopic_diseases = ifelse(family_Atopic_diseases_sum >0,1,0),
                    mother_Atopic_diseases = ifelse(mother_Atopic_diseases_sum>0,1,0),
                    father_Atopic_diseases = ifelse(father_Atopic_diseases_sum > 0, 1,0),
                    sibling_Atopic_diseases = ifelse(sibling_Atopic_diseases_sum > 0 ,1, 0))

# Changing to binary variables:
df$gender[df$gender == 2] <- 0
df$`Jew/Arab`[df$`Jew/Arab` == 2] <- 0
df$food_allergy_24_month[df$food_allergy_24_month == 2] <- 0
df$inhalation_24_month[df$inhalation_24_month == 2] <- 0
df$prevention_therapy[df$prevention_therapy == 2] <- 0
df$AD_24_month[df$AD_24_month == 2] <- 0
df$steroids_treatment24[df$steroids_treatment24 == 2] <- 0
df$ABX_treatment_24_months[df$ABX_treatment_24_months == 2] <- 0
df$hospitalization_24_months[df$hospitalization_24_months == 2] <- 0
df$chronic_disease_24[df$chronic_disease_24 == 2] <- 0
df$type_of_pregnancy[df$type_of_pregnancy == 2] <- 0

#Changing NA to 0 for 
df$number_ABX_treatment_24_months<- replace(df$number_ABX_treatment_24_months, is.na(df$number_ABX_treatment_24_months), 0)
df$hospitalization_24_months<- replace(df$hospitalization_24_months, is.na(df$hospitalization_24_months), 0)

df$prevention_therapy<- replace(df$prevention_therapy, is.na(df$prevention_therapy), 0)
df$number_inhalation<- replace(df$number_inhalation, is.na(df$number_inhalation), 0)

df$AD_treatment_24_month<- replace(df$AD_treatment_24_month, is.na(df$AD_treatment_24_month), 0)
df$steroids_treatment24<- replace(df$steroids_treatment24, is.na(df$steroids_treatment24), 0)
df$number_steroid_treat<- replace(df$number_steroid_treat
, is.na(df$number_steroid_treat), 0)

df$number_of_ABX_first_year<- replace(df$number_of_ABX_first_year, is.na(df$number_of_ABX_first_year), 0)




df <- df %>% drop_na(AD_first_year, FA_first_year, ABX_fisrt_year,food_allergy_24_month, inhalation_24_month, AD_24_month) 
```


```{r}
#Creating dependent variables:
#option A of explanatory var:
df <- df %>% mutate(is_atopic_first_year = ifelse((AD_first_year + FA_first_year + ABX_fisrt_year)>0,1,0), 
                    is_atopic_second_year = ifelse((food_allergy_24_month + inhalation_24_month + AD_24_month)>0,1,0))
#option B of explanatory var:
#From mail
```

```{r}
#Choosing explanatory variables: Ones that have a sufficient number of observation before and after Corona
#The distribution of observations according to the timeline of the corona. Before the corona = 1, the first six months of the corona = 2  and after = 3:

group1 <- as.Date("2019-03-01")
group2 <- as.Date("2020-03-11")
group3 <- as.Date("2021-02-07")
df <- df %>% mutate(covid_time = ifelse(DOB < group1, 1, ifelse(DOB < group2, 2, ifelse(DOB<group3, 3, 4))))
table(df$covid_time)

# at this moment there is no sutible data from group 4 - post corona 
attach(df)
```
# Summary of the data:


```{r}
# Frequency table for each column
tablelist <- mapply(table, df %>% select(-c(record_id, DOB, mother_year_birth)))
# tablelist
```

```{r}
# How many NA are in each column
NA_count <- as.data.frame(t(df %>% summarise_all(funs(sum(is.na(.)))))) %>% dplyr::rename("amount of NA" = 1) 
NA_count <- NA_count %>% mutate(Name = rownames(NA_count))
NA_column = NA_count %>% filter(`amount of NA` >0) %>% arrange(`amount of NA`)
```


# arrange all the colnames by type
```{r}
# MASBIR
full_binari_exp <- c('Jew/Arab', 'gender', 'Allergy_total', 
                'mother_asthma', 'mother_atopic_dermatitis', 'mother_Allergic_rhinitis', 'mother_food_allergy', 
                'father_asthma', 'father_atopic_dermatitis', 'father_Allergic_rhinitis', 'father_food_allergy', 
                'sibling1_asthma', 'sibling1_atopic_dermatitis', 'sibling1_Allergic_rhinitis', 'sibling1_food_allergy', 
                'sibling2_asthma', 'sibling2_atopic_dermatitis', 'sibling2_Allergic_rhinitis', 'sibling2_food_allergy', 
                'sibling3_asthma', 'sibling3_atopic_dermatitis', 'sibling3_Allergic_rhinitis', 'sibling3_food_allergy', 
                'sibling4_asthma', 'sibling4_atopic_dermatitis', 'sibling4_Allergic_rhinitis', 'sibling4_food_allergy', 
                'vaccine_12mo', 'prevention_therapy', 'steroids_treatment24',
                'hospitalization_24_months', 'chronic_disease_24',
                'family_Atopic_diseases', 'mother_Atopic_diseases', 'father_Atopic_diseases', 'sibling_Atopic_diseases')

binari_exp <- c('Jew/Arab', 'gender', 'Allergy_total', 
                'mother_asthma', 'mother_atopic_dermatitis', 'mother_Allergic_rhinitis', 'mother_food_allergy', 
                'father_asthma', 'father_atopic_dermatitis', 'father_Allergic_rhinitis', 'father_food_allergy', 
                'vaccine_12mo', 'food_allergy_24_month', 'inhalation_24_month', 'prevention_therapy', 'steroids_treatment24',
                'ABX_treatment_24_months', 'hospitalization_24_months', 'chronic_disease_24',
                'family_Atopic_diseases', 'mother_Atopic_diseases', 'father_Atopic_diseases', 'sibling_Atopic_diseases')

continues_exp <- c('mother_year_birth', 'DOB', 'mother_education')

discrite_exp <- c('SES', 'number_of_children', 'preganancy_number', 'number_of_ABX_first_year', 'number_inhalation', 'number_steroid_treat', 'number_ABX_treatment_24_months',
                  'family_Atopic_diseases_sum', 'mother_Atopic_diseases_sum', 'father_Atopic_diseases_sum', 'sibling_Atopic_diseases_sum', 
                  'sibling_asthma_sum', 'sibling_AD_sum', 'sibling_Allergic_rhinitis_sum', 'sibling_FA_sum')

categoric_exp <- c('type_of_pregnancy', 'nursing_setup_at_12mo', 'daycare_24month', 'vaccines_24month', 'AD_treatment_24_month', 'chronic_rhinitis_24_month', 'covid_time')

#MOOSBAR
exploreatory_variables <- c('HRAD_first_year', 'AD_first_year', 'FA_first_year', 'ABX_fisrt_year',
                            'inhalation_24_month', 'AD_24_month', 'food_allergy_24_month', 'ABX_treatment_24_months',
                            'is_atopic_first_year', 'is_atopic_second_year') 

```


# Two dim table - covid time VS each column 
```{r}
covid_time_table <- function(kpi){
  paste(kpi)
  return(table(df$covid_time, df[[kpi]]))}

# sapply(binari_exp, covid_time_table, simplify =FALSE)
# sapply(discrite_exp, covid_time_table, simplify =FALSE)
# sapply(categoric_exp, covid_time_table, simplify =FALSE)
# sapply(exploreatory_variables, covid_time_table, simplify =FALSE)
```


plots for presentation:
# is atopic vs is atopic and corona
```{r,fig.width=5}

plot_df_1 <- df %>% select(covid_time, AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% group_by(covid_time) %>% summarise(AD_first_year = sum(AD_first_year), AD_24_month = sum(AD_24_month), FA_first_year = sum(FA_first_year),food_allergy_24_month = sum(food_allergy_24_month), ABX_fisrt_year = sum(ABX_fisrt_year),inhalation_24_month = sum(inhalation_24_month)) %>% gather(chronic_disease, count, AD_first_year:inhalation_24_month, factor_key=TRUE)

plot_df_2 <- df %>% select(covid_time,is_atopic_first_year,is_atopic_second_year) %>% group_by(covid_time) %>% summarise(is_atopic_first_year= sum(is_atopic_first_year), is_atopic_second_year= sum(is_atopic_second_year)) %>% gather(chronic_disease, count, is_atopic_first_year:is_atopic_second_year, factor_key=TRUE)

ggplot(plot_df_1,aes(x=chronic_disease, y= count, fill = as.factor(covid_time))) +
    geom_col() + xlab("Chronic Disease") + ylab("Cases") +
  scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'),name = "Group")

ggplot(plot_df_2,aes(x=chronic_disease, y= count, fill = as.factor(covid_time))) +
    geom_col(position = "dodge")+ xlab("Chronic Disease") + ylab("Cases") +
  scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'),name = "Group")# + scale_fill_discrete(name = "Group")
```
#SES vs is atopic
```{r}
# ggplot(data=df, aes(x=SES, group= as.factor(is_atopic_first_year), fill=as.factor(is_atopic_first_year))) +
#     geom_density(adjust=1.5, alpha=.4) 
# 
# ggplot(data=df, aes(x=SES, group= as.factor(AD_first_year), fill=as.factor(AD_first_year))) +
#     geom_density(adjust=1.5, alpha=.4) 
# 
# ggplot(data=df, aes(x=SES, group= as.factor(is_atopic_second_year), fill=as.factor(is_atopic_second_year))) +
#     geom_density(adjust=1.5, alpha=.4) 
```

```{r}
plot_df_3 <- df %>% select(SES,is_atopic_first_year,is_atopic_second_year) %>% drop_na() %>% group_by(SES) %>% summarise(is_atopic_first_year= sum(is_atopic_first_year), is_atopic_second_year= sum(is_atopic_second_year)) %>% gather(chronic_disease, count, is_atopic_first_year:is_atopic_second_year, factor_key=TRUE)

plot_df_3 <- df %>% select(SES,is_atopic_first_year,is_atopic_second_year) %>% drop_na() %>% group_by(SES) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year)) %>% gather(chronic_disease, count, is_atopic_first_year:is_atopic_second_year, factor_key=TRUE)

ggplot(plot_df_3,aes(x=as.factor(SES), y= count, fill = as.factor(chronic_disease))) +
    geom_col(position = "dodge") + scale_fill_discrete(name = "Group")+ xlab("SES") + ylab("Cases") +
  scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

table(SES)
```

#Age of mother vs is atopic
```{r}
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(is_atopic_first_year), fill=as.factor(is_atopic_first_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(is_atopic_second_year), fill=as.factor(is_atopic_second_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(AD_first_year), fill=as.factor(AD_first_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(AD_24_month), fill=as.factor(AD_24_month))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(FA_first_year), fill=as.factor(FA_first_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(food_allergy_24_month), fill=as.factor(food_allergy_24_month))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(ABX_fisrt_year), fill=as.factor(ABX_fisrt_year))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(inhalation_24_month), fill=as.factor(inhalation_24_month))) +
#     geom_density(adjust=1.5, alpha=.4)
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(is_atopic_first_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(is_atopic_second_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(AD_first_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(AD_24_month), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(FA_first_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(food_allergy_24_month), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(ABX_fisrt_year), y=mother_age)) + 
#   geom_boxplot()
# 
# ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=as.factor(inhalation_24_month), y=mother_age)) + 
#   geom_boxplot()
# 
```


#mother_education va is atopic
```{r,fig.width=10}
# plot_df_4 <- df %>% select(mother_education,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(mother_education) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_4,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(mother_education))) +
#       geom_col(position = "dodge") + scale_fill_discrete(name = "mother_education")+ xlab("mother_education") + ylab("percent diseas") #+scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))
```

#number_of_children vs is atopic
```{r,fig.width=10}
# plot_df_5 <- df %>% select(number_of_children,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(number_of_children) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_5,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(number_of_children))) +
#     geom_col(position = "dodge") + scale_fill_discrete(name = "number_of_children")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))
```

#preganancy_number vs is atopic
```{r,fig.width=10}
# plot_df_6 <- df %>% select(preganancy_number,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(preganancy_number) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_6,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(preganancy_number))) +
#     geom_col(position = "dodge") + scale_fill_discrete(name = "preganancy_number")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

```

# type_of_pregnancy vs ia atopic
```{r,fig.width=7}
# plot_df_7 <- df %>% select(type_of_pregnancy,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(type_of_pregnancy) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)
# 
# ggplot(plot_df_7,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(type_of_pregnancy))) +
#     geom_col(position = "dodge") + scale_fill_discrete(name = "type_of_pregnancy")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))
```


#number_of_ABX_first_year vs is atopic
```{r,fig.width=6}
plot_df_88 <- df %>% select(number_of_ABX_first_year,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(number_of_ABX_first_year) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)

ggplot(plot_df_88,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(number_of_ABX_first_year))) +
    geom_col(position = "dodge") + scale_fill_discrete(name = "number_of_ABX_first_year")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

```

# number_ABX_treatment_24_months vs is atopic
```{r,fig.width=12}
plot_df_8 <- df %>% select(number_ABX_treatment_24_months,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(number_ABX_treatment_24_months) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)

ggplot(plot_df_8,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(number_ABX_treatment_24_months))) +
    geom_col(position = "dodge") + scale_fill_discrete(name = "number_ABX_treatment_24_months")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

```

#nursing_setup_at_12mo vs is atopic
```{r,fig.width=12}
plot_df_10 <- df %>% select(nursing_setup_at_12mo,is_atopic_first_year,is_atopic_second_year,AD_first_year, AD_24_month, FA_first_year,food_allergy_24_month, ABX_fisrt_year,inhalation_24_month) %>% drop_na() %>% group_by(nursing_setup_at_12mo) %>% summarise(is_atopic_first_year= mean(is_atopic_first_year), is_atopic_second_year= mean(is_atopic_second_year),AD_first_year = mean(AD_first_year), AD_24_month = mean(AD_24_month), FA_first_year = mean(FA_first_year),food_allergy_24_month = mean(food_allergy_24_month), ABX_fisrt_year = mean(ABX_fisrt_year),inhalation_24_month = mean(inhalation_24_month)) %>% gather(chronic_disease, mean, is_atopic_first_year:inhalation_24_month, factor_key=TRUE)

ggplot(plot_df_10,aes(x=as.factor(chronic_disease), y= mean, fill = as.factor(nursing_setup_at_12mo))) +
    geom_col(position = "dodge") + scale_fill_discrete(name = "nursing_setup_at_12mo")+ xlab("chronic_disease") + ylab("percent of disease")# +scale_fill_manual(values=c('dodgerblue4', 'darkorange1', 'green4'))

```

#binary vs catagoric:
```{r, warning=F}
cat <- c("SES","mother_education", "number_of_children","preganancy_number","type_of_pregnancy","number_of_ABX_first_year","nursing_setup_at_12mo")

binary = c("father_food_allergy","HRAD_first_year","family_Atopic_diseases","mother_Atopic_diseases","father_Atopic_diseases","sibling_Atopic_diseases")


for (a in cat) {
  for (b in binary) {
    a_index <- grep(paste0("^",a,"$") , colnames(df))
    b_index <- grep(paste0("^",b,"$"), colnames(df))
    x <- sort(unique(unlist(df[,a_index])))
    y <- sort(unique(unlist(df[,b_index])))
    data <- expand.grid(X=x, Y=y)
    data$Z <- c(table(unlist(df[,a_index]),unlist(df[,b_index])))
    p = ggplot(data, aes(as.factor(X), as.factor(Y), fill= Z)) +
      geom_tile()+
      scale_fill_gradient(low="red", high="blue") + xlab(a) + ylab(b)
    print(p)
  }
}

gg <- function(kpi){ list(percents = round(ftable(df[[kpi]], family_Atopic_diseases) / rowSums(ftable(df[[kpi]], family_Atopic_diseases)),3), n = ftable(df[[kpi]], family_Atopic_diseases))}
sapply(cat, gg,  simplify =FALSE)
```


#binary vs continues:
```{r, fig.width=6}
for (c in binary) {
  c_index <- grep(paste0("^",c,"$"), colnames(df))
  p =ggplot(data=df%>% mutate(mother_age = 2023 - mother_year_birth), aes(x=mother_age, group= as.factor(unlist(df[,c_index])), fill=as.factor(unlist(df[,c_index])))) +
    geom_density(adjust=1.5, alpha=.4)+ scale_fill_discrete(name = c)
  print(p)
  
}
```


# continues exploratoty variables - DOB Vs Atopic deasese 

```{r, warning=F}
exloretory_variables = list('AD_first_year', 'AD_24_month', 'FA_first_year', 'food_allergy_24_month', 'ABX_fisrt_year', 'inhalation_24_month', 'is_atopic_first_year', 'is_atopic_second_year')

DOB_explanatory <- function(kpi){
  return (ggplot(df, aes(x=DOB, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_density(adjust=1.5, alpha=.4) + ggtitle(kpi))}
lapply(exloretory_variables, DOB_explanatory)
```


# Continues exploratoty VS Categoty exploratoty
```{r, warning=F}
# Categoric_variables <- c('SES', 'Jew/Arab', 'gender', 'study_group_COMEET', 'mother_education', 'number_of_children', 'preganancy_number', 'type_of_pregnancy', "family_Atopic_diseases", "mother_Atopic_diseases", "father_Atopic_diseases", "sibling_Atopic_diseases" )
# 
# DOB_explanatory_dist <- function(kpi){
#   return (ggplot(df, aes(x=DOB, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_density(adjust=1.5, alpha=.4) + ggtitle(kpi))}
# lapply(Categoric_variables , DOB_explanatory_dist)
# 
# DOB_explanatory_dist <- function(kpi){
#   return (ggplot(df, aes(x=DOB, group = as.factor(df[[kpi]]), fill=as.factor(df[[kpi]]))) + geom_boxplot(adjust=1.5, alpha=.4) + ggtitle(kpi))}
# lapply(Categoric_variables , DOB_explanatory_dist)
```


# Categoty exploratoty VS Categoty exploratoty
```{r, warning=F}
# # categoric_exp
# # discrite_exp
# 
# family <- function(kpi){
#   return (ggplot(df, aes(family_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , family)
# 
# mother <- function(kpi){
#   return (ggplot(df, aes(mother_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , mother)
# 
# father <- function(kpi){
#   return (ggplot(df, aes(father_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , father)
# 
# sibling <- function(kpi){
#   return (ggplot(df, aes(sibling_Atopic_diseases_sum, fill=as.factor(df[[kpi]]))) + geom_bar(adjust=1.5) + ggtitle(paste0(kpi, ' vs. family_Atopic_diseases')))}
# lapply(categoric_exp , sibling)
```




# Binary VS binary

```{r}
library(corrplot)
df_binary_first<- df %>% select("Jew/Arab", "gender", "type_of_pregnancy", "family_Atopic_diseases", "mother_Atopic_diseases","father_Atopic_diseases", "sibling_Atopic_diseases", "HRAD_first_year", "AD_first_year", "FA_first_year", "ABX_fisrt_year", "is_atopic_first_year")

df_binary_second<- df %>% select("Jew/Arab", "gender", "type_of_pregnancy", "family_Atopic_diseases", "mother_Atopic_diseases","father_Atopic_diseases", "sibling_Atopic_diseases", "vaccine_12mo", "prevention_therapy", "vaccines_24month", "steroids_treatment24", "hospitalization_24_months", "chronic_disease_24", "food_allergy_24_month", "inhalation_24_month", "AD_24_month", "is_atopic_second_year")

cor_1<- cor(df_binary_first, method = c("spearman"), use="pairwise.complete.obs")
corrplot(cor_1, method = 'color', order = 'FPC', type = 'lower', diag = FALSE, col = COL2('BrBG'))

cor_2<- cor(df_binary_second, method = c("spearman"), use="pairwise.complete.obs")
corrplot(cor_2, method = 'color', order = 'FPC', type = 'lower', diag = FALSE, col=COL2('BrBG'))
```









